Perceptions and Preferences for Meritocracy

### 2020-04-12 docs/paper.Rmd
update paper

b. Beliefs: the "get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy. Nevertheless, another version of this same battery asks about "how important you think it _should_ be". Which one of both is the "belief"? If we consider belief as related to a normative content, actually the more close version is the "should be", being the other one more close to the perception of the situation. Such distinction was already made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy, whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”. Is the belief in meritocracy a perception or a preference? A comprehensive measurement should be in line with what @duru-bellat_whos_2012 did, in order to open up possibilities of analyzing whether perceptions and preferences are actually the same (i.e. correlation close to 1) or they are different aspects of the same phenomenon.

c. Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocracti aspects as talent, letting this out of the measurement. A different approach was followed by @kunovich_systems_2007, who decide to include non-meritocratic elements in the measurement of meritocracy. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar aproach was taken by @newman_false_2015, reverse-coding non-meritocratic items. The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic. Nevertheless, despite making this distinction the author ends up substracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.

@@

- Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
- Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
- Based on previous indicators as far as possible, for the sake of comparability between studies
- Brief, as to be used in a regular survey.

@@

The first one distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy. Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The second axis consider the distinction between meritocratic and non-meritocratic dimensions. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them some adaptations of previous items that where streamlined in order to a better fit with the specific dimensions. For meritocratic indicators we use effort and talent as the main components of the traditional concept as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.

@@

* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.
* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.
* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.
* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.

>>>>

b. Beliefs: the "get ahead" battery refers to "how important you think it is", considered by Mijs as a belief in meritocracy. Nevertheless, another version of this same battery asks about "how important you think it _should_ be". Which one of both is the "belief"? If we consider belief as related to a normative content, actually the more close version is the "should be", being the other one more close to the perception of the situation. Such distinction was already made by @duru-bellat_whos_2012, who used the item "how important should the number of years spent in education and training be in deciding how much money people ought to earn?" for "desired" meritocracy, whereas for "perceived" meritocracy they use two items: “Would you say that in your country, people are rewarded for their efforts?” and “... people are rewarded for their skills?”. Is the belief in meritocracy a perception or a preference? A comprehensive measurement should be in line with what @duru-bellat_whos_2012 did, in order to open up possibilities of analyzing whether perceptions and preferences are actually the same (i.e. correlation close to 1) or they are different aspects of the same phenomenon. As @son_hing_merit_2011-1 have pointed out, "People can believe that outcomes ought to be distributed on the basis of merit and yet vary in their perceptions of whether this is how society currently operates" (p. 435). Accordingly, we suggest that the use of the term belief is confusing, being more accurate to speak about meritocratic perceptions and preferences.

c. Non-meritocratic aspects: @mijs_paradox_2019 make reference to some non-meritocratic aspects as talent, letting this out of the measurement. A different approach was followed by @kunovich_systems_2007, who decide to include non-meritocratic elements in the measurement of meritocracy. Using the items’ battery listing a number of reasons about “How important should be in deciding pay...” (as @duru-bellat_whos_2012 for desired meritocracy), he decided that reasons as education and responsibility are meritocratic and pointed 1 if considered essential, whereas reasons such as having a family and children were pointed 1 if they were considered "not important at all" (i.e. reverse coded). A similar approach was taken by @newman_false_2015, reverse-coding non-meritocratic items, similarly to what occurs with the "Preference for the Merit Principle Scale" [@davey_preference_1999]. The assumption that meritocratic and non-meritocratic elements are the poles of the same continuum was analyzed by @reynolds_perceptions_2014 using the same "get ahead" perceptions' battery items. They consider education, ambition and hard work as meritocratic and other reasons such as wealthy family and right connections and non-mertitocratic. Nevertheless, despite making this distinction the author ends up substracting one dimension from the other, assuming that they are two poles of the same continuum as @kunovich_systems_2007 did.

@@

- Multidimensional, incorporating previous distinctions between preferences and perceptions as well as between meritocratic and non-meritocratic aspects.
- Multiple indicators for each dimension, in order to account for measurement error in a confirmatory factor analysis context.
- Based on previous indicators as far as possible, for the sake of comparability between studies
- Brief, as to be used in a regular survey. In this point it differs for instance from the proposal of the "Preference for the Merit Principle Scale" [@davey_preference_1999], as they use 15 items just for one dimension (besides the problem of reverse-coding non-meritocratic items).

@@

The first axis distinguishes between different types of "beliefs", using instead the terms perceptions and preferences for meritocracy [@duru-bellat_whos_2012; @son_hing_merit_2011-1]. Perceptions refers to the extent to which people see meritocracy working in their society, which in terms of measurement relates to the "reasons to get ahead" battery, whereas preferences refer to normative expectations that are usually linked to a "should" expression (e.g. whether hard work should be related to payment). The second axis consider the distinction between meritocratic and non-meritocratic dimensions [@reynolds_perceptions_2014]. This aspect has been usually treated as different ends of a same continuum in part of the previous research, an assumption that requieres empirical scrutiny. These non-meritocratic elements usually refer to the use of personal contacts or family advantages to get ahead in life.

Regarding the selection of indicators, most of them some adaptations of previous items that where streamlined in order to a better fit with the specific dimensions. For meritocratic indicators we use effort and talent as the main components of the traditional concept of merit as defined by @young_rise_1962, whereas for non-meritocratic dimensions we use having rich parents and good contacts.

@@

* H1. The perception of meritocracy is a latent variable based on indicators of the importance attributed to talent and the effort to get ahead in life.

* H2. The non-meritocratic perception is a latent variable that derives from two indicators related to the agreement with the statement that people with contacts and rich parents manage to get ahead.

* H3. Meritocratic preferences behave as the latent variable based on a normative value of effort and talent.

* H4. Non-meritocratic preferences behave as a latent variable based on the normative value of the use of personal contacts and having rich parents.
